Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2021-09-08 paper/paper.tex
address comments

\section{Generic programming}\label{sec:generic}
\red{I suggest we have a section on generic programming, putting both the semiring stuff and the summary table to this section, since generic programming is one of the highlights of this paper as well. I think it doesn't need to be long. Jinguo, could you write one or two paragraphs introducing generic programming?}

\red{Generic programming is ...}
\red{dispatch to different types etc. ...}

In previous works combining tensor network and combinatoric problems~\cite{},
the elements in tensor networks are limited to regular number types such as floating point numbers and integers.
As the development of modern compiling technology, we do not need to limit our imaginations to regular number types anymore.
The key technology is called the generic programming:

@@

This definition of generic programming contains two major pieces, one is a single program for the general setting, and another is performance.
To understand the first point, suppose we want to write a function to take the power of an element $f(x, n) := x^n$,
the generic programming approach does not assume $x$ being a regular number,
instead it treats $x$ as an element with an associative multiplication operation $\odot$ and multiplicative identity $\mymathbb{1}$.
In such a way, when the program takes a matrix as input, it computes the matrix power without rewriting this function for matrices.
The second point about performance basically rules out dynamic typed languages like Python,
because in these languages, the efficiency of a generic type is not guaranteed.
Imagine the speed of computing the matrix multiplication between two numpy arrays with python objects as elements.
The generic programming that we want to discuss is those in static types languages like C++, and Julia~\cite{Bezanson2012}.
C++ uses templates for generic programming while Julia has just in time compiling and multiple dispatch.
When the program ``sees'' a new input type, the compiler recompiles the generic program for the new type.
A lot optimization can be done during the compiling, like inlining immutable elements with fixed sizes in an array.
In Julia, these inlined arrays can even be compiled to GPU devices for faster computation~\cite{Besard2018}.

This leads us to think about what is the most general element type that allowed in a tensor network contraction program.
We find as long as the tensor elements form a commutative semiring, the tensor network contraction result can be contraction order independent.
A semiring is a ring without additive inverse, while a commutative semiring is a semiring where the multiplication operation is commutative.
To define a commutative semiring with the addition operation $\oplus$ and the multiplication operation $\odot$ on a set $R$, the following relations must hold for arbitrary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b \oplus a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} \odot a = \mymathbb{0}
\end{align*}
The requirement of being commutative is for the tensor contraction result to be independent of the contraction order, which could be relaxed in certain cases. In the following sections, we show how to compute the independence polynomial,
the maximal independence polynomial, independence number, the number of independent sets, and enumeration of the MIS configurations of a general graph $G$ by designing tensor element types as commutative semirings while keeping the tensor network generic~\cite{Stepanov2014}.
Table~\ref{tbl:generictypes} summarizes the problems that can be solved by various tensor element types.

@@

\begin{example}\label{eg:tensorcontraction}
Mapping a graph (left) to a tensor network (right) that encodes the independence polynomial. In the generalized tensor network's graphical representation, a vertex is mapped to a hyperedge, and an edge is mapped to an edge tensor. \red{a vertex is mapped to a hyperedge?}

@@

Before contracting the tensor network and evaluating the independence polynomial numerically, let us first elevate the tensor elements $0$s and $1$s in tensors $W(x)$ and $B$ from integers and floating point numbers to the additive identity,
$\mymathbb{0}$, and multiplicative identity, $\mymathbb{1}$, of a commutative semiring as discussed in Sec.~\ref{sec:generic}.
The most natural approach is to treat the tensor elements as polynomials and evaluate the polynomial directly.
Let us create a polynomial type, and represent a polynomial $a_0 + a_1 x + \ldots + a_k x^k$ as a coefficient vector $(a_0, a_1, \ldots, a_k) \in \mathbb{R}^k$, so, e.g., $x$ is represented as $(0, 1)$.
We define the algebra between the polynomials $a$ of order $k_a$ and $b$ of order $k_b$ as
\begin{equation}
    \begin{split}
    a \oplus b &= (a_0 + b_0, a_1 + b_1, \ldots, a_{\max(k_a, k_b)} + b_{\max(k_a, k_b)}),\\
    a \odot b &= (a_0 + b_0, a_1b_0 + a_0b_1, a_{2}b_{0} + a_{1}b_{1} + a_{0}b_{2},  \ldots, a_{k_a} b_{k_b}),\\
    \mymathbb{0} &= (), \text{\red{should this be (0)?}} \\
    \mymathbb{1} &= (1).\label{eq:polynomial}
    \end{split}
\end{equation}
We can see these operations are standard addition and multiplication operations of polynomials, and the polynomial type forms a commutative ring. The tensors $W$ and $B$ can thus be written as 
\begin{equation}
    W_{s_i}^{\rm poly} = \left(\begin{matrix}
        \mymathbb{1} \\
        (0,1)
    \end{matrix}\right),   
    \qquad \qquad
        B_{s_i s_j}^{\rm poly} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
By contracting the tensor network with the polynomial type, the final result is the exact representation of the independence polynomial.
One way to efficiently evaluate the multiplication operation is to use the convolution theorem~\cite{Schonhage1971}, but this approach suffers from a space overhead proportional to $\alpha(G)$ because each polynomial requires a vector of such size to store the coefficients.

Here, we propose to find the independence polynomial by fitting $\alpha(G)+1$ random values of $x_{i}$ and $y_{i} = I(G,x_{i})$. One can then compute the independence polynomial coefficients $a_{i}$ by solving the linear equation: 
\begin{equation}
\left(\begin{matrix}
1 & x_0 & x_0^2 & \ldots & x_0^{\alpha(G)} \\
1 & x_1 & x_1^2 & \ldots & x_1^{\alpha(G)} \\
\vdots & \vdots & \vdots &\ddots & \vdots \\
1 & x_{\alpha(G)} & x_{\alpha(G)}^2 & \ldots & x_{\alpha(G)}^{\alpha(G)}
\end{matrix}\right)
\left(\begin{matrix}
a_0 \\ a_1 \\ \vdots \\ a_{\alpha(G)}
\end{matrix}\right)
= \left(\begin{matrix}
y_0 \\ y_1 \\ \vdots \\ y_{\alpha(G)}
\end{matrix}\right).\label{eq:lineareq}
\end{equation}
With this approach, we do not incur the linear overhead in space. However, because the independence polynomial coefficients can have a huge order-of-magnitude range, if we use floating point numbers in the computation, the round-off errors can be significant for the counting of large independent set sizes.
In addition, the number could easily overflow if we use fixed-width integer types.
The big integer type is also not a good option because big integers with varying width can be very slow and is incompatible with graphics processing unit (GPU) devices. These problems can be solved by introducing a finite-field algebra $\text{GF}(p)$:
\begin{equation}
\begin{split}
    x ~\oplus~ y &= x+y\pmod p,\\
    x ~\odot~ y &= xy\pmod p,\\
    \mymathbb{0} &= 0,\\
    \mymathbb{1} &= 1.
\end{split}\label{eq:finitefield}
\end{equation}
With a finite-field algebra, we have the following observations:
\begin{enumerate}
    \item One can use Gaussian elimination~\cite{Golub2013} to solve the linear equation \Eq{eq:lineareq} since it is a generic algorithm that works for any elements with field algebra. The multiplicative inverse of a finite-field algebra can be computed with the extended Euclidean algorithm.
    \item Given the remainders of a larger unknown integer $x$ over a set of co-prime integers $\{p_1, p_2, \ldots, p_n\}$,
    $x \pmod {p_1 \times p_2 \times \ldots \times p_n}$ can be computed using the Chinese remainder theorem. With this, one can infer big integers from small integers.
\end{enumerate}
With these observations, we develop Algorithm~\ref{alg:finitefield} to compute the independence polynomial exactly without introducing space overheads. In the algorithm, except the computation using the Chinese remainder theorem, all computations are done with integers of fixed width $W$.
In \App{app:fft}, we provide another method to solve the linear equation using discrete Fourier transform.

@@

        compute the tensor contraction on $\text{GF}(p)$ and obtain $Y = (y_0, y_1, \ldots , y_{\alpha(G)}) \pmod p $\; \red{tensor network contraction for each $i$}

@@

        \If{$A_P = A_{P \times p}$}{
            \Return $A_P$ \tcp*[l]{converged}
        }
        $P = P \times p$\;
    }\caption{Exactly compute the independence polynomial of a graph $G$ without integer overflow}\label{alg:finitefield} \red{discuss this algorithm table, $X_{ij}$, etc.}
\end{algorithm}

\subsection{Maximal independence polynomial}
\red{It might be helpful to add a figure of the tensor network for computing the maximal independence polynomial.} 
Sometimes people are interested in knowing maximal solutions to understand why their programs are trapped in a local minimal.
In this paper, the word ``maximal'' is different with the ``maximum'' discussed the previous discussion in such a way that it is not necessarily refer to the gloal optimum.
Instead of counting all independent sets, the maximal independence polynomial counts the number of maximal independent sets of various sizes~\cite{Hu2017}.
 Concretely, it is defined as
\begin{equation}
I_{\rm max}(G, x) = \sum_{k=0}^{\alpha(G)} b_k x^k,
\end{equation}
where $b_k$ is the number of maximal independent sets of size $k$ in $G$.
Obviously, $b_{k} \leq a_{k}$ and $b_{\alpha(G)} = a_{\alpha(G)}$. $I_{\rm max}(G, 1)$ counts the total number of maximal independent sets~\cite{Gaspers2012, Manne2013}, where the fastest algorithm currently has a runtime of $O(1.3642^{|V|})$~\cite{Gaspers2012}.
For the problem of finding the MIS, $b_{k}$ counts the number of local optimum at size $k < \alpha(G)$, and can, in some cases, provide hints on the difficulty of finding the MIS using local algorithms~\red{cite experiment}.
The uni-modality, log-concavity, and real-rootness properties of the maximal independence polynomial for special classes of graphs have also been studied~\cite{Hu2017}.

@@

With the tensors defined as $T(x_v)$, we can perform a similar computation of contracting the tensor network with the same tensor element types as described in the previous section~\ref{sec:indpoly}, the result of which then produces the maximal independence polynomial.
The computational complexity of this new tensor network contraction is often greater than the one for computing the independence polynomial.
However, in most sparse graphs, this tensor network contraction approach is still significantly faster than enumerating all the maximal cliques on its complement graph using the Bron-Kerbosch algorithm~\cite{Bron1973}, which is the standard algorithm that we are aware of to compute the maximal independence polynomial.

\section{Maximum independent sets and its counting problem}
\subsection{Tropical algebra for finding the independence number and counting MISs}
In the previous section, we focused on computing the independence polynomial for a graph $G$ of given independence number $\alpha(G)$, but we did not show how to compute this number.
The method we use to compute this quantity is based on the following observations. Let $x=\infty$, the independence polynomial becomes
\begin{equation}
I(G, \infty) = a_{\alpha(G)} \infty^{\alpha(G)},
\end{equation}
where the lower-order terms vanish. We can thus replace the polynomial type $a = (a_0, a_1, \ldots, a_k)$ with $a(k) = (a_{k}, k)$, where the second element stores the largest exponent $k$ and the first element stores the corresponding coefficient $a_{k}$.
From this, we can define a new algebra as \red{I've changed this, but I don't have strong feelings about this. We can certainly consider changing back.
I thought this would be more consistent with the earlier notations and show how it's stored in the program as well.}
\begin{equation}
\begin{split}
    a(k) \oplus a(j) &= \begin{cases}
        \left(a_k + a_j, \max(k,j) \right), & k = j \\
        \left(a_j, \max(k,j) \right), & k < j \\
        \left(a_k, \max(k,j) \right), & k > j
    \end{cases}, \\
    a(k) \odot a(j) &= (a_k a_j, k+j) \\
    \mymathbb{0} &= (0, -\infty)\\
    \mymathbb{1} &= (1, 0). \label{eq:countingtropical}
\end{split}
\end{equation}
%\begin{equation}
%\begin{split}
%    a_x\infty^x \oplus a_y\infty^y &= \begin{cases}
%        (a_x + a_y)\infty^{\max(x,y)}, & x = y\\
%        a_y\infty^{\max(x,y)}, & x < y\\
%        a_x\infty^{\max(x,y)}, & x > y
%    \end{cases}, \\
%    a_x\infty^x \odot a_y\infty^y &= a_x a_y\infty^{x+y}\\
%    \mymathbb{0} &= 0\infty^{-\infty}\\
%    \mymathbb{1} &= 1\infty^{0}. \label{eq:countingtropical}
%\end{split}
%\end{equation}
% In the program, we only need to store the exponent $x$ and the corresponding coefficient $a_x$ initialized to $1$. 
The algebra of the exponents becomes the max-plus tropical algebra: $k \oplus j = \max(k,j)$ and $k \odot j = k + j$~\cite{Maclagan2015, Moore2011}.
This algebra is the same as the one used in Liu et al.~\cite{Liu2021} to calculate and count spin glass ground states.
For independent set calculations here, the vertex tensor and edge tensor becomes:
\begin{equation}
    W_{s_i}^{\rm tropical} = \left(\begin{matrix}
        \mymathbb{1} \\
        (1,1)
    \end{matrix}\right),   
    \qquad \qquad
        B_{s_i s_j}^{\rm tropical} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

\subsection{Truncated polynomial algebra for counting independent sets of large size}
Instead of counting just the MISs, one may be interested in counting the independent sets of large sizes close to the MIS size.
For example, if one is interested in counting only $a_{\alpha(G)}$ and $a_{\alpha(G)-1}$, we can define a truncated polynomial algebra by keeping only the largest two coefficients in the polynomial in \Eq{eq:polynomial}:
\begin{equation}
    \begin{split}
    a \oplus b &= (a_{\max(k_a, k_b)-1} + b_{\max(k_a, k_b)-1}, a_{\max(k_a, k_b)} + b_{\max(k_a, k_b)}),\\
    a \odot b &= (a_{k_a-1} b_{k_b}+a_{k_a} b_{k_b-1}, a_{k_a} b_{k_b}),\\
    \mymathbb{0} &= (), \text{\red{should this be (0)?}}  \\
    \mymathbb{1} &= (1).\label{eq:max2poly}
    \end{split}
\end{equation}
In the program, we thus need a data structure that contains three fields, the largest order $k$, and the coefficients for the two largest orders $a_k$ and $a_{k-1}$.
This approach can clearly be extended to calculate more independence polynomial coefficients and is more efficient than calculating the entire independence polynomial.
\red{can be used to enumerate suboptimal solution as well?}

@@

Let us rearrange the coefficients $r^j$ to $a_j$, the matrix on the left side becomes the discrete Fourier transform matrix. Thus, we can obtain the coefficients by inverse Fourier transform $\vec a_r = {\rm FFT^{-1}}(\omega) \cdot \vec y$, where $(\vec a_r)_j = a_j r ^j$.
By choosing different $r$, one can obtain better precision in low independent set size region  ($\omega<1$) or high independent set size region ($\omega>1$). \red{is that $\omega < 1$ or $r <1$?}

>>>>

\section{Generic programming}\label{sec:generic}
\red{I suggest we have a section on generic programming, putting both the semiring stuff and the summary table to this section, since generic programming is one of the highlights of this paper as well. I think it doesn't need to be long. Jinguo, could you write one or two paragraphs introducing generic programming?}\blue{I wrote a draft here, does it look ok in general?}
In previous works combining tensor network and combinatoric problems~\cite{Kourtis2019,Biamonte2017},
the elements in tensor networks are limited to regular number types such as floating point numbers and integers.
As the development of modern compiling technology, we do not need to limit our imaginations to regular number types anymore.
The key technology playing the role is called the generic programming:

@@

This definition of generic programming contains two major pieces, one is a single program for the general setting, and another is about efficiency.
To understand the first point, suppose we want to write a function to take the power of an element $f(x, n) := x^n$.
One can easily write a function for regular numbers that computes the power of $x$ in $\log(n)$ steps using the multiply and square trick.
A generic programmer does not assume $x$ being a regular number type,
instead it treats $x$ as an element with an associative multiplication operation $\odot$ and multiplicative identity $\mymathbb{1}$.
In such a way, when the program takes a matrix as input, it computes the matrix power without extra efforts.
The second point about performance. For dynamic typed languages like Python,
one can easily write very general codes, however, the efficiency is not guaranteed.
Imagine the speed of computing the matrix multiplication between two numpy arrays with python objects as elements.
The generic programming that we want to discuss is those in static types languages like C++, and Julia~\cite{Bezanson2012}.
C++ uses templates for generic programming while Julia has just in time compiling and multiple dispatch.
When they ``sees'' a new input type, the compiler recompiles the generic program for the new type.
A lot optimization can be done during the compiling, like inlining immutable elements with fixed sizes in an array to decrease the cache miss rate when accessing data.
In Julia, these inlined arrays can even be compiled to GPU devices for faster computation~\cite{Besard2018}.

This leads us to think about what is the most general element type that allowed in a tensor network contraction program.
Our finding is: as long as the tensor elements form a commutative semiring, the tensor network contraction result can be contraction order independent.
So what is a commutative semiring?
A semiring is a ring without additive inverse, while a commutative semiring is a semiring where the multiplication operation is commutative.
To define a commutative semiring with the addition operation $\oplus$ and the multiplication operation $\odot$ on a set $R$, the following relations must hold for arbitrary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b \oplus a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} \odot a = \mymathbb{0}
\end{align*}
The requirement of being commutative is for the tensor contraction result to be independent of the contraction order, which could be relaxed in certain cases. In the following sections, we show how to compute the independence polynomial,
the maximal independence polynomial, independence number, the number of independent sets, and enumeration of the MIS configurations of a general graph $G$ by designing tensor element types as commutative semirings while keeping the tensor network generic~\cite{Stepanov2014}.
Table~\ref{tbl:generictypes} summarizes the problems that can be solved by various tensor element types.

@@

\begin{example}\label{eg:tensorcontraction}
Mapping a graph (left) to a tensor network (right) that encodes the independence polynomial. In the generalized tensor network's graphical representation, a vertex is mapped to a hyperedge, and an edge is mapped to an edge tensor.
\red{a vertex is mapped to a hyperedge?}\blue{$\checkmark$}

@@

Before contracting the tensor network and evaluating the independence polynomial numerically, let us first elevate the tensor elements $0$s and $1$s in tensors $W(x)$ and $B$ from integers and floating point numbers to the additive identity,
$\mymathbb{0}$, and multiplicative identity, $\mymathbb{1}$, of a commutative semiring as discussed in Sec.~\ref{sec:generic}.
The most natural approach is to treat the tensor elements as polynomials and evaluate the polynomial directly.
Let us create a polynomial type, and represent a polynomial $a_0 + a_1 x + \ldots + a_k x^k$ as a coefficient vector $(a_0, a_1, \ldots, a_k) \in \mathbb{R}^k$, so, e.g., $x$ is represented as $(0, 1)$.
We define the algebra between the polynomials $a$ of order $k_a$ and $b$ of order $k_b$ as
\begin{equation}
    \begin{split}
    a \oplus b &= (a_0 + b_0, a_1 + b_1, \ldots, a_{\max(k_a, k_b)} + b_{\max(k_a, k_b)}),\\
    a \odot b &= (a_0 + b_0, a_1b_0 + a_0b_1, a_{2}b_{0} + a_{1}b_{1} + a_{0}b_{2},  \ldots, a_{k_a} b_{k_b}),\\
    \mymathbb{0} &= (), \text{\red{should this be (0)?}}\blue{\text{both are ok}} \\
    \mymathbb{1} &= (1).\label{eq:polynomial}
    \end{split}
\end{equation}
We can see these operations are standard addition and multiplication operations of polynomials, and the polynomial type forms a commutative ring. The tensors $W$ and $B$ can thus be written as 
\begin{equation}
    W_{s_i}^{\rm poly} = \left(\begin{matrix}
        \mymathbb{1} \\
        (0,1)
    \end{matrix}\right),   
    \qquad \qquad
        B_{s_i s_j}^{\rm poly} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
By contracting the tensor network with the polynomial type, the final result is the exact representation of the independence polynomial.
One way to efficiently evaluate the multiplication operation is to use the convolution theorem~\cite{Schonhage1971}, but this approach suffers from a space overhead proportional to $\alpha(G)$ because each polynomial requires a vector of such size to store the coefficients.

Here, we propose to find the independence polynomial by fitting $\alpha(G)+1$ random values of $x_{i}$ and $y_{i} = I(G,x_{i})$. One can then compute the independence polynomial coefficients $a_{i}$ by solving the linear equation: 
\begin{equation}
\left(\begin{matrix}
1 & x_0 & x_0^2 & \ldots & x_0^{\alpha(G)} \\
1 & x_1 & x_1^2 & \ldots & x_1^{\alpha(G)} \\
\vdots & \vdots & \vdots &\ddots & \vdots \\
1 & x_{\alpha(G)} & x_{\alpha(G)}^2 & \ldots & x_{\alpha(G)}^{\alpha(G)}
\end{matrix}\right)
\left(\begin{matrix}
a_0 \\ a_1 \\ \vdots \\ a_{\alpha(G)}
\end{matrix}\right)
= \left(\begin{matrix}
y_0 \\ y_1 \\ \vdots \\ y_{\alpha(G)}
\end{matrix}\right).\label{eq:lineareq}
\end{equation}
With this approach, we do not incur the linear overhead in space. However, because the independence polynomial coefficients can have a huge order-of-magnitude range, if we use floating point numbers in the computation, the round-off errors can be significant for the counting of large independent set sizes.
In addition, the number could easily overflow if we use fixed-width integer types.
The big integer type is also not a good option because big integers with varying width can be very slow and is incompatible with graphics processing unit (GPU) devices. These problems can be solved by introducing a finite-field algebra $\text{GF}(p)$:
\begin{equation}
\begin{split}
    x ~\oplus~ y &= x+y\pmod p,\\
    x ~\odot~ y &= xy\pmod p,\\
    \mymathbb{0} &= 0,\\
    \mymathbb{1} &= 1.
\end{split}\label{eq:finitefield}
\end{equation}
With a finite-field algebra, we have the following observations:
\begin{enumerate}
    \item One can use Gaussian elimination~\cite{Golub2013} to solve the linear equation \Eq{eq:lineareq} since it is a generic algorithm that works for any elements with field algebra. The multiplicative inverse of a finite-field algebra can be computed with the extended Euclidean algorithm.
    \item Given the remainders of a larger unknown integer $x$ over a set of co-prime integers $\{p_1, p_2, \ldots, p_n\}$,
    $x \pmod {p_1 \times p_2 \times \ldots \times p_n}$ can be computed using the Chinese remainder theorem. With this, one can infer big integers from small integers.
\end{enumerate}
With these observations, we develop Algorithm~\ref{alg:finitefield} to compute the independence polynomial exactly without introducing space overheads.
This is an iterative algorithm that tries different prime numbers $p$ until convergence.
In each iteration, we contract tensor networks to evaluate the polynomial for each $x\in X$ on the finite field algebra ${\rm GF}(p)$.
Let use denote the results as $Y$, then we solve a linear equation $Y = \hat X A_p$ using the gaussian elimination on ${\rm GF}(p)$ to find the coefficient moduluses $A_p \equiv A \pmod p$.
Finally we apply the Chinese remainder theorem to get $A \pmod P$, where $P$ is a product of all prime numbers chosen in current and previous iterations.
If this number does not change comparing with the previous iteration, it indicates a convergent result.
All computations are done with integers of fixed width $W$ except the last step of applying Chinese remainder theorem.
In \App{app:fft}, we provide another method to solve the linear equation using discrete Fourier transform.

@@

        compute the tensor contraction for each $x = X_i \pmod p$ on $\text{GF}(p)$ and obtain $Y = (y_0, y_1, \ldots , y_{\alpha(G)}) \pmod p $\;

@@

        \If{$A_P = A_{P \times p}$}{
            \Return $A_P$ \tcp*[l]{converged}
        }
        $P = P \times p$\;
    }\caption{Compute the independence polynomial exactly without integer overflow}\label{alg:finitefield} \red{discuss this algorithm table, $X_{ij}$, etc.}\blue{done}
\end{algorithm}

\subsection{Maximal independence polynomial}
\red{It might be helpful to add a figure of the tensor network for computing the maximal independence polynomial.} \blue{done}
Sometimes people are interested in knowing maximal solutions to understand why their programs are trapped in a local minimal.
In this paper, the word ``maximal'' is different with the ``maximum'' discussed the previous discussion in such a way that it is not necessarily refer to the gloal optimum.
Instead of counting all independent sets, the maximal independence polynomial counts the number of maximal independent sets of various sizes~\cite{Hu2017}.
 Concretely, it is defined as
\begin{equation}
I_{\rm max}(G, x) = \sum_{k=0}^{\alpha(G)} b_k x^k,
\end{equation}
where $b_k$ is the number of maximal independent sets of size $k$ in $G$.
Obviously, $b_{k} \leq a_{k}$ and $b_{\alpha(G)} = a_{\alpha(G)}$. $I_{\rm max}(G, 1)$ counts the total number of maximal independent sets~\cite{Gaspers2012, Manne2013}, where the fastest algorithm currently has a runtime of $O(1.3642^{|V|})$~\cite{Gaspers2012}.
For the problem of finding the MIS, $b_{k}$ counts the number of local optimum at size $k < \alpha(G)$, and can, in some cases, provide hints on the difficulty of finding the MIS using local algorithms~\red{cite experiment}.
The uni-modality, log-concavity, and real-rootness properties of the maximal independence polynomial for special classes of graphs have also been studied~\cite{Hu2017}.

@@

With the tensors defined as $T(x_v)$, we can perform a similar computation of contracting the tensor network with the same tensor element types as described in the previous section~\ref{sec:indpoly}, the result of which then produces the maximal independence polynomial.

    \centerline{\begin{tikzpicture}[
    dot/.style = {circle, fill, minimum size=#1,
                inner sep=0pt, outer sep=0pt},
    dot/.default = 6pt  % size of the circle diameter 
                    ]  
        \def\dx{0};
        \def\r{0.5cm}
        \def\G{1.0}
        \foreach \x/\y/\v in {0/0/a, 1/0/b, 2/0/c, 3/0/d, 4/0/e}
            \node[color=black] at (\x*\G+\dx,\y) (\v) {$s_\v$};
        \foreach \x/\v/\t in {0/A/$T_a$, 1/B/$T_b$, 2/C/$T_c$, 3/D/$T_d$, 4/E/$T_e$}
            \node[color=white,fill=black,dot=\r] at (\x*\G+\dx,1.5) (\v) {\t};
        \draw [cyan,thick] (a) -- (A);
        \draw [cyan,thick] (a) -- (B);
        \draw [cyan,thick] (a) -- (C);
        \draw [blue,thick] (b) -- (B);
        \draw [blue,thick] (b) -- (A);
        \draw [blue,thick] (b) -- (C);
        \draw [blue,thick] (b) -- (D);
        \draw [red,thick] (c) -- (C);
        \draw [red,thick] (c) -- (A);
        \draw [red,thick] (c) -- (B);
        \draw [red,thick] (c) -- (D);
        \draw [green,thick] (d) -- (D);
        \draw [green,thick] (d) -- (B);
        \draw [green,thick] (d) -- (C);
        \draw [green,thick] (d) -- (E);
        \draw [orange,thick] (e) -- (E);
        \draw [orange,thick] (e) -- (D);
    \end{tikzpicture}}

One can see the average degree of a tensor is increase.
The computational complexity of this new tensor network contraction is often greater than the one for computing the independence polynomial.
However, in most sparse graphs, this tensor network contraction approach is still significantly faster than enumerating all the maximal cliques on its complement graph using the Bron-Kerbosch algorithm~\cite{Bron1973}, which is the standard algorithm that we are aware of to compute the maximal independence polynomial.
Let us consider the example in \Sec{eg:tensorcontraction}, its corresponding tensor network structure for computing maximal independent sets now becomes:

\section{Maximum independent sets and its counting problem}
\subsection{Tropical algebra for finding the independence number and counting MISs}
In the previous section, we focused on computing the independence polynomial for a graph $G$ of given independence number $\alpha(G)$, but we did not show how to compute this number.
The method we use to compute this quantity is based on the following observations. Let $x=\infty$, the independence polynomial becomes
\begin{equation}
I(G, \infty) = a_{\alpha(G)} \infty^{\alpha(G)},
\end{equation}
where the lower-order terms vanish. We can thus replace the polynomial type $a = (a_0, a_1, \ldots, a_k)$ with $a(k) = (a_{k}, k)$, where the second element stores the largest exponent $k$ and the first element stores the corresponding coefficient $a_{k}$.
From this, we can define a new algebra as \red{I've changed this, but I don't have strong feelings about this. We can certainly consider changing back.
I thought this would be more consistent with the earlier notations and show how it's stored in the program as well.}
\blue{I prefer the previous one, what about you @XG?}
\begin{equation}
\begin{split}
    a(k) \oplus a(j) &= \begin{cases}
        \left(a_k + a_j, \max(k,j) \right), & k = j \\
        \left(a_j, \max(k,j) \right), & k < j \\
        \left(a_k, \max(k,j) \right), & k > j
    \end{cases}, \\
    a(k) \odot a(j) &= (a_k a_j, k+j) \\
    \mymathbb{0} &= (0, -\infty)\\
    \mymathbb{1} &= (1, 0). \label{eq:countingtropical}
\end{split}
\end{equation}
%\begin{equation}
%\begin{split}
%    a_x\infty^x \oplus a_y\infty^y &= \begin{cases}
%        (a_x + a_y)\infty^{\max(x,y)}, & x = y\\
%        a_y\infty^{\max(x,y)}, & x < y\\
%        a_x\infty^{\max(x,y)}, & x > y
%    \end{cases}, \\
%    a_x\infty^x \odot a_y\infty^y &= a_x a_y\infty^{x+y}\\
%    \mymathbb{0} &= 0\infty^{-\infty}\\
%    \mymathbb{1} &= 1\infty^{0}. \label{eq:countingtropical}
%\end{split}
%\end{equation}
% In the program, we only need to store the exponent $x$ and the corresponding coefficient $a_x$ initialized to $1$. 
The algebra of the exponents becomes the max-plus tropical algebra: $k \oplus j = \max(k,j)$ and $k \odot j = k + j$~\cite{Maclagan2015, Moore2011}.
This algebra is the same as the one used in Liu et al.~\cite{Liu2021} to calculate and count spin glass ground states.
For independent set calculations here, the vertex tensor and edge tensor becomes:
\begin{equation}
    W_{s_i}^{\rm tropical} = \left(\begin{matrix}
        \mymathbb{1} \\
        (1,1)
    \end{matrix}\right),   
    \qquad \qquad
        B_{s_i s_j}^{\rm tropical} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

\subsection{Truncated polynomial algebra for counting independent sets of large size}
Instead of counting just the MISs, one may be interested in counting the independent sets of large sizes close to the MIS size.
For example, if one is interested in counting only $a_{\alpha(G)}$ and $a_{\alpha(G)-1}$, we can define a truncated polynomial algebra by keeping only the largest two coefficients in the polynomial in \Eq{eq:polynomial}:
\begin{equation}
    \begin{split}
    a \oplus b &= (a_{\max(k_a, k_b)-1} + b_{\max(k_a, k_b)-1}, a_{\max(k_a, k_b)} + b_{\max(k_a, k_b)}),\\
    a \odot b &= (a_{k_a-1} b_{k_b}+a_{k_a} b_{k_b-1}, a_{k_a} b_{k_b}),\\
    \mymathbb{0} &= (), \text{\red{should this be (0)?}} \blue{\text{both are fine}} \\
    \mymathbb{1} &= (1).\label{eq:max2poly}
    \end{split}
\end{equation}
In the program, we thus need a data structure that contains three fields, the largest order $k$, and the coefficients for the two largest orders $a_k$ and $a_{k-1}$.
This approach can clearly be extended to calculate more independence polynomial coefficients and is more efficient than calculating the entire independence polynomial.
As will be shown bellow, this algebra can also be extended to enumerate suboptimal solutions.
\red{can be used to enumerate suboptimal solution as well?}
\blue{added some description}

@@

Let us rearrange the coefficients $r^j$ to $a_j$, the matrix on the left side becomes the discrete Fourier transform matrix. Thus, we can obtain the coefficients by inverse Fourier transform $\vec a_r = {\rm FFT^{-1}}(\omega) \cdot \vec y$, where $(\vec a_r)_j = a_j r ^j$.
By choosing different $r$, one can obtain better precision in low independent set size region  ($r<1$) or high independent set size region ($r>1$). \red{is that $\omega < 1$ or $r <1$?}\blue{it is $r$, fixed}
